Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-11-08 Manuscript/manuscript.tex

\todo{Samuli: I think that we either have to do the analysis in SI properly,
  i.e., using the real $C_I$ from simulations or remove the sentence below and
the related concent from SI.\\
Joe: I think it shows that ternary complex makes a good fit to our simulation data. 
The fact that we use a different (and more accurate) estimate of $C_I$ and other quantities does not hamper this statement.}

@@

\todo{This should be either done properly with the correct $C_I$ or removed. Removing
might be a better idea at this point.}
It was found in the original work~\cite{altenbach84} that 
a ternary complex binding model (i.e.~2~POPC:1~\ce{Ca^{2+}})
provides the best fit to experimental measurements of all considered models in that study. 
In such a model, there is a linear relationship between quantities 
$C_b$, mole fraction of bound \ce{Ca^{2+}} per POPC, and $\sqrt{C_b/C_I}$, 
where $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
The concentration $C_b$ was obtained from an extrapolation of linear relation 
between deuterium NMR measurements and atomic absorption spectroscopy for low concetrations of \ce{CaCl2}.
Such an extrapolation is valid as long as the mode of \ce{Ca^{2+}} binding 
remains constant throughout the extrapolation range. 
The concentration $C_I$ is determined by using the surface potential by using the Boltzmann equation.
However, Boltzmann theory yields inaccurate results
for divalent cations like \ce{Ca^{2+}}~\cite{Andelman1995}. 
An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions.
Hence we hypothesise that the discrepancy between the results in the experiment~\cite{altenbach84} and 
our simulations likely lays in the fact that 
the assumptions and relations used for determinig concentrations $C_b$ and $C_I$ in the experiment~\cite{altenbach84}
gradually do not hold for higher concentrations of \ce{Ca^{2+}}.

\begin{figure}[tbp]
  \centering
  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
  \caption{\label{fig:cacl-bind}
    Ternary complex binding model of \ce{Ca^{2+}} to a POPC membrane 
    that assumes the stoichiometry of 2~POPC:1~\ce{Ca^{2+}} (details in reference~\citenum{altenbach84}) 
    provides a good fit to experimental measurements~\cite{altenbach84}
    and it also provides a good fit to our simulation data. 
    Note that the units in the reference~\citenum{altenbach84} are different from the units presented here,
    and, hence, the observed slope of the linear relationship is different.
    }
\end{figure}

>>>>

@@

It was found in the original work~\cite{altenbach84} that 
a ternary complex binding model (i.e.~2~POPC:1~\ce{Ca^{2+}})
provides the best fit to experimental measurements of all considered models in that study. 
In such a model, there is a linear relationship between quantities 
$C_b$, mole fraction of bound \ce{Ca^{2+}} per POPC, and $\sqrt{C_b/C_I}$, 
where $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
The concentration $C_b$ was obtained from an extrapolation of linear relation 
between deuterium NMR measurements and atomic absorption spectroscopy for low concetrations of \ce{CaCl2}.
Such an extrapolation is valid as long as the mode of \ce{Ca^{2+}} binding 
remains constant throughout the extrapolation range. 
The concentration $C_I$ is determined from the surface potential using the Boltzmann equation.
However, Boltzmann theory yields inaccurate results
for divalent cations like \ce{Ca^{2+}}~\cite{Andelman1995}. 
An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions;
$C_b$ can be calculated from simulation using the robust definition in section~\ref{sec:affinity},
and $C_I$ can be obtained from the density profiles in Fig.~\ref{fig:cacl-density} 
as the minimum concentration at the shallow depression separating the bulk water phase 
and the peak of adsorbed cations around 2.6\,nm from the membrene centre. 
The small discrepancy between the results in the experiment~\cite{altenbach84} and 
our simulations are likely caused by the difference in the evaluation of the concentrations.

\begin{figure}[tbp]
  \centering
  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
  \caption{\label{fig:cacl-bind}
    Ternary complex binding model of \ce{Ca^{2+}} to a POPC membrane 
    that assumes the stoichiometry of 2~POPC:1~\ce{Ca^{2+}} (details in reference~\citenum{altenbach84}) 
    provides a good fit to experimental measurements~\cite{altenbach84}
    and it also provides a good fit to our simulation data. 
    This supports our statement that the success of the ternary complex model in the experiments
    can be understood with the atomistic detail of our simulations as an average 
    stoichiometry of otherwise almost evenly distributed clusters of one cation with 1,2 or 3 lipids (Fig.~\ref{fig:cacl_complexes}). 
    Note that the units in the reference~\citenum{altenbach84} are different from the units presented here,
    and, hence, the observed slope of the linear relationship is slightly different.
    }
\end{figure}
